DeepSphere: a graph-based spherical CNN

### 2019-11-13 main.tex
edit main body for new theorem

\paragraph{Theoretical convergence.}
We extend the work of Belkin and Nyiogi to a sufficiently regular, deterministic sampling scheme. For this purpose we need now to introduce some definitions. Following their setting, we define the extended graph Laplacian operator as the linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\begin{equation}
	\label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=0}^{n-1} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
	\end{equation}
and the rescaled graph Laplacian operator
    \begin{equation}
	\label{eq:rescaled Heat Kernel Graph Laplacian operator}
	 \hat L_n^t := \frac{|\mathbb S^2|}{4\pi t^2}L_n^t
	\end{equation}
Note that we have the following relations $\b{L}_n^t \b{f} = T_\V L_n^t f$, $\b{\hat L} = \frac{|\mathbb S^2|}{4\pi t^2}\b L_n^t$. Given a sampling scheme $\V = \{x_0, \dots, x_{n-1}\}$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ as the largest distance between the center $x_i$ and any point on the surface $\sigma_i$.
 Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.
Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere
\begin{equation}\label{eq:rotation operator}
	\mathcal R_g f(x) := f(g^{-1}(x)).
\end{equation}
\begin{theorem}
	For a sampling $\V = \{x_i\in\S^2\}_{i=0}^{n-1}$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that 
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$  
	\begin{equation*}
		\lim_{n\to\infty}\hat L_n^{t_n}f(y) = \triangle_{\S^2}f(y).
	\end{equation*} 
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, the discrete counterpart $ \mathbf{\hat L}$ converges pointwise towards the Laplace-Beltrami operator ($\Delta_{\S^2}$) in the sense explained in \cite{belkin2007convergence}. This operator commutes with rotation and hence theorem \ref{theo:pointwise convergence for a regular sampling} is a strong hint for equivariance, as formalized in the following proposition.
%%%%%%%%%%%%%%%%%%%%%%%  link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{prop}\label{prop:equivariance}
Under the hypothesis of theorem \ref{theo:pointwise convergence for a regular sampling}, the rescaled graph Laplacian commutes with any rotation, in the limit of infinite sampling, i.e.,
\begin{equation*}
\forall y\in\mathbb S^2\quad\left| \mathcal R_g \hat L_n^{t_n} f (y) - \hat L_n^{t_n} \mathcal R_g f(y) \right| \xrightarrow{n\to\infty}0.
\end{equation*}
\end{prop}
%%%%%%%%%%%%%%%%%%%%%%% end of link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%

Importantly, the proof of Theorem \ref{theo:pointwise convergence for a regular sampling} in Appendix \ref{sec: appendix: proof of theorem} inspires us in the construction of our graph Laplacian. It is inspirational for the graph construction and for how the parameter $t$ needs to be adjusted with respect to the number of points $n$.

Nevertheless, it is important to keep in mind the limits of Theorem \ref{theo:pointwise convergence for a regular sampling}.
First, it is \emph{not} a proof of equivariance of the graph convolution for a specific sampling. To construct such a proof, a stronger convergence result is needed, for example spectral convergence (instead of pointwise).
% \todo{Can we intuitively write what pointwise convergence means, and how it is more limited than spectral or uniform convergence?}
We note that in the random sampling case, the graph laplacian does \emph{not} converge uniformly in general \cite[Sec. 3]{belkin2007convergence}, but does for a class of Lipschitz functions \cite[Sec. 6]{belkin2005towards}.  Second, while do not have a formal proof for it, we strongly believe that the HEALPix sampling does satisfy the hypothesis $d^{(n)}\leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$ with $\alpha$ very close or equal to $\linefrac{1}{2}$. The empirical results discussed in the next paragraph also points in this direction. This is further discussed in Appendix \ref{sec: appendix: proof of theorem}.
% 3) Show experimental results
\paragraph{Empirical convergence.}

@@

\begin{figure}

	\begin{minipage}{0.6\linewidth}
		\centering
		\includegraphics[width=1\linewidth]{all_in_one.pdf}

		\caption{Normalized mean equivariance error \eqref{eq:mean equivariance error}. The number of pixels is $n = 4 s^2$ for the equiangular sampling and $n = 12 s^2$ for the HEALPix sampling.}
		\label{fig:equivariance_error}
	\end{minipage}
	\hfill
	\begin{minipage}{0.35\linewidth}
		\centering
		\includegraphics[width=\linewidth]{kernel_widths}
		\caption{Optimal kernel widths.}
		\label{fig:kernel_widths}
		\vspace{1em}
		\includegraphics[height=4.5em]{lamp_000018}
		\hfill
		\includegraphics[height=4.5em]{lamp_000018_sphere_nobar}
		\caption{3D object represented as a spherical depth map.}
		\label{fig:depthmap}
		\vspace{1em}
		\includegraphics[width=\linewidth]{spectrum}
		\caption{Power spectral densities.}
		\label{fig:spectrum}
	\end{minipage}
	\vspace{-0.5cm}
\end{figure}

@@

\section{Proof of proposition \ref{prop:equivariance}}
\begin{proof}
Fix $x\in \mathbb S^2$. Since any rotation $\mathcal R_g$ is an isometry, and the Laplacian $\Delta$ commutes with all isometries of a Riemanniann manifold, and defining $\mathcal R_g f =: f'$ for ease of notation, we can write that 
	\begin{align*}
			\seminorm{\Rg \Ln f (x) - \Ln \Rg f(x) } &\leq \seminorm{\Rg \Ln f (x) - \Rg \laplacian f(x)} + \seminorm{\Rg \laplacian f (x) - \Ln \Rg f(x)} =\\
			& =\seminorm{\Rg (\Ln f - \laplacian f ) (x)} + \seminorm{ \laplacian f' (x) - \Ln f'(x)} \leq \\
			& \leq \seminorm{(\Ln f - \laplacian f ) (g^{-1}(x))} + \seminorm{ \laplacian f' (x) - \Ln f'(x)}
	\end{align*}

>>>>

\paragraph{Theoretical convergence.}
We extend the work of \citep{belkin2005towards} to a sufficiently regular, deterministic sampling scheme. 
Following their setting, we will work with the extended graph Laplacian operator as the linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\begin{equation}
	\label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=0}^{n-1} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
	\end{equation}
This operator extends the graph Laplacian operator previously defined with the weight scheme of \eqref{eqn:weights:belkin} to each point of the manifold, i.e. the sphere. 
As the radius of the kernel $t$ will be adapted to the number of sample, we also rescale the graph Laplacian operator as
    % \begin{equation}
	% \label{eq:rescaled Heat Kernel Graph Laplacian operator}
	 $\hat{L}_n^t := \frac{|\mathbb S^2|}{4\pi t^2}L_n^t$
	% \end{equation}
Hence, we have the following relations $\b{L}_n^t \b{f} = T_\V L_n^t f$. 
% $\b{\hat L} = \frac{|\mathbb S^2|}{4\pi t^2}\b L_n^t$. \nati{why define this, we are not using it}
Given a sampling scheme $\V = \{x_0, \dots, x_{n-1}\}$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ as the largest distance between the center $x_i$ and any point on the surface $\sigma_i$.
 Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.
Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere
\begin{equation}\label{eq:rotation operator}
	\mathcal R_g f(x) := f(g^{-1}(x)).
\end{equation}
\begin{theorem}
	For a sampling $\V = \{x_i\in\S^2\}_{i=0}^{n-1}$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that 
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$  
	\begin{equation*}
		\lim_{n\to\infty}\hat{L}_n^{t_n}f(y) = \triangle_{\S^2}f(y).
	\end{equation*} 
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, the discrete counterpart of $ \mathbf{\hat L}$ converges pointwise towards the Laplace-Beltrami operator ($\Delta_{\S^2}$) in the sense explained in \cite{belkin2007convergence}. This operator commutes with rotation and hence Theorem \ref{theo:pointwise convergence for a regular sampling} is a strong step toward equivariance. We formalize in the following proposition.
%%%%%%%%%%%%%%%%%%%%%%%  link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{theorem}\label{theo:equivariance}
Under the hypothesis of theorem \ref{theo:pointwise convergence for a regular sampling}, the rescaled graph Laplacian commutes with any rotation, in the limit of infinite sampling, i.e.,
\begin{equation*}
\forall y\in\mathbb S^2\quad\left| \mathcal R_g \hat L_n^{t_n} f (y) - \hat L_n^{t_n} \mathcal R_g f(y) \right| \xrightarrow{n\to\infty}0.
\end{equation*}
\end{theorem}
%%%%%%%%%%%%%%%%%%%%%%% end of link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%

Importantly, one of the important benefit of Theorem \ref{theo:pointwise convergence for a regular sampling} is that its proof (in Appendix \ref{sec: appendix: proof of theorem}) inspires us in the construction of our graph Laplacian. 
In particular, it gives us insides for how the parameter $t$ needs to be adjusted with respect to the number of points $n$.

% Nevertheless, it is important to keep in mind the limits of Theorem \ref{theo:pointwise convergence for a regular sampling}.
% First, it is \emph{not} a proof of equivariance of the graph convolution for a specific sampling. To construct such a proof, a stronger convergence result is needed, for example spectral convergence (instead of pointwise).
% \todo{Can we intuitively write what pointwise convergence means, and how it is more limited than spectral or uniform convergence?}
We note that in the random sampling case, the graph laplacian does \emph{not} converge uniformly in general \cite[Sec. 3]{belkin2007convergence}, but does for a class of Lipschitz functions \cite[Sec. 6]{belkin2005towards}.  
Second, while do not have a formal proof for it, we strongly believe that the HEALPix sampling does satisfy the hypothesis $d^{(n)}\leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$ with $\alpha$ very close or equal to $\linefrac{1}{2}$. The empirical results discussed in the next paragraph also points in this direction. This is further discussed in Appendix \ref{sec: appendix: proof of theorem}.
% 3) Show experimental results

\paragraph{Empirical convergence.}

@@

\begin{figure}
	\begin{minipage}{0.6\linewidth}
		\centering
		\includegraphics[width=1\linewidth]{all_in_one.pdf}

		\caption{Normalized mean equivariance error \eqref{eq:mean equivariance error}. The number of pixels is $n = 4 s^2$ for the equiangular sampling and $n = 12 s^2$ for the HEALPix sampling.}
		\label{fig:equivariance_error}
	\end{minipage}
	\hfill
	\begin{minipage}{0.35\linewidth}
		\centering
		\includegraphics[width=\linewidth]{kernel_widths}
		\caption{Optimal kernel widths.}
		\label{fig:kernel_widths}
		\vspace{1em}
		\includegraphics[height=4.5em]{lamp_000018}
		\hfill
		\includegraphics[height=4.5em]{lamp_000018_sphere_nobar}
		\caption{3D object represented as a spherical depth map.}
		\label{fig:depthmap}
		\vspace{1em}
		\includegraphics[width=\linewidth]{spectrum}
		\caption{Power spectral densities.}
		\label{fig:spectrum}
	\end{minipage}
\end{figure}

@@

\section{Proof of Theorem~\ref{theo:equivariance}}
\begin{proof}
Fix $x\in \mathbb S^2$. Since any rotation $\mathcal R_g$ is an isometry, and the Laplacian $\Delta$ commutes with all isometries of a Riemanniann manifold, and defining $\mathcal R_g f =: f'$ for ease of notation, we can write that 
	\begin{align*}
			\seminorm{\Rg \Ln f (x) - \Ln \Rg f(x) } &\leq \seminorm{\Rg \Ln f (x) - \Rg \laplacian f(x)} + \seminorm{\Rg \laplacian f (x) - \Ln \Rg f(x)} =\\
			& =\seminorm{\Rg (\Ln f - \laplacian f ) (x)} + \seminorm{ \laplacian f' (x) - \Ln f'(x)} \leq \\
			& \leq \seminorm{(\Ln f - \laplacian f ) (g^{-1}(x))} + \seminorm{ \laplacian f' (x) - \Ln f'(x)}
	\end{align*}
